Sequential algorithms to split and merge large 3D images

### 2017-05-31 paper.tex
Added models

% packages
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{hyperref}

@@

\section{Algorithms}

@@

@@

Although simplistic, Algorithms~\ref{algo:naive-slices}
and~\ref{algo:naive-blocks} actually exhibit very different
performance due to the amount of seeks performed by Algorithm 1.b. The
number of seeks performed in Algo 1.b for each block is indeed the
number of rows in a slice, that is, $(F/N)^{1/3}$. The total number of
seeks, considering $N$ blocks is thus $N^{2/3}.F^{1/3}$. In practice,
this could lead to a tremendous slowdown.

@@

\subsection{Cluster reads}

@@

\subsection{Multiple reads}

@@

@@

\subsection{Implementation and Benchmark}

As a running example to benchmark our algorithms, we use the 76~GB
BigBrain image split in 125 chunks. The data is available [here]. We
perform experiments using the two disks below:
- HD ...
- SSD ...
Our algorithms are implemented in Python using the Nibabel library.

@@

-- Splitting, 125 blocks, HD vs SSD.
Figure 1: on HD, 5 algorithms, 10 repetitions per algorithm.
Figure 2: on SSD, 5 algorithms, 10 repetitions per algorithm.

-- Merging, 125 blocks, HD vs SSD.
Figure 3: on HD, 5 algorithms, 10 repetitions per algorithm.
Figure 4: on SSD, 5 algorithms, 10 repetitions per algorithm.

-- HD, 250 blocks vs 125 blocks. 
Figure 5: different block sizes

-- HD, 125 blocks, F vs F/2
- Figure 6: Different images sizes, with constant memory.

>>>>

% packages
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{hyperref}
\usepackage{mathtools}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

@@

\section{Algorithms and Models}

@@

\subsection{Notations}

We adopt the following notations:
\begin{itemize}
\item $R$: number of voxels in the reconstructed images.
\item $b$: bytes per voxel in the reconstructed images and splits (in B).
\item $n$: number of splits (blocks or slices).
\item $\alpha_r$: data read rate from disk (in B/s).
\item $\alpha_w$: data write rate to disk (in B/s).
\item $\beta$: disk access time (in s).
\item $m$: amount of RAM available (in B).
\end{itemize}

The following properties are straightforward applications of these
definitions:
\begin{itemize}
\item bytes per chunk: $\frac{Rb}{n}$
\item number of chunks that can fit in memory: $\frac{m.n}{Rb}$
\end{itemize}
For the sake of mathematical simplicity, the models below assume that
the reconstructed image and blocks are cubic. Under this assumption,
we have the following definitions and properties:
\begin{itemize}
\item Number of slices in a block: $\sqrt[3]{\frac{R}{n}}$
\item Number of blocks required to fully cover a slice in the
  reconstructed image: $\sqrt[3]{n}^2=n^{\frac{2}{3}}$
  \item Number of full slices that can be reconstructed from $k$
    blocks: $\floor*{\frac{k}{n^{\frac{2}{3}}}}\sqrt[3]{\frac{R}{n}}$
\end{itemize}
\todo{Make a figure to illustrate the notations.}

@@

Although simplistic, Algorithms~\ref{algo:naive-slices}
and~\ref{algo:naive-blocks} actually exhibit very different
performance due to the amount of seeks performed by Algorithm 1.b. The
number of seeks performed in Algo 1.b for each block is indeed the
number of rows in a slice, that is, $(R/n)^{1/3}$. The total number of
seeks, considering $n$ blocks is thus $N^{2/3}.F^{1/3}$. In practice,
this could lead to a tremendous slowdown.

For slices, the total merge time is expressed as follows:
\begin{equation}
T_\mathrm{slices} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w}+2\beta n
\end{equation}
and for blocks:
\begin{equation}
  T_\mathrm{blocks} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w}+\beta n\left(1+\sqrt[3]{\frac{R}{n}}\right)
\end{equation}

@@

\subsection{Buffered slices}

\begin{equation}
T_\mathrm{buff\_slices} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w}+\beta\left( n + \ceil*{\frac{bR}{m}}\right)
\end{equation}

\subsection{Buffered blocks: Cluster reads}

@@

    \begin{equation}
      \begin{multlined}
  T_\mathrm{cluster\_reads} = \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w} +\\
  \beta
  \left[n+ 
  \ceil*{\frac{bR}{m}} \left[
           \sqrt[3]{\frac{R}{n}} \left(
              \ceil*{\frac{m\sqrt[3]{n}}{bR}} - \floor*{\frac{m\sqrt[3]{n}}{bR}}
           \right)+1
             \right]
             \right]
             \end{multlined}
\end{equation}

\subsection{Buffered blocks: Multiple reads}

@@

     \begin{equation}
       T_\mathrm{multiple\_reads}= \frac{bR}{\alpha_r}+\frac{bR}{\alpha_w} + \beta\ceil*{\frac{bR}{m}}\left(1+ \ceil*{\frac{m.n}{bR}}\right)\\
\end{equation}

@@

\subsection{Benchmark}
\label{sec:benchmark}

\subsubsection{Data}
As a running example to benchmark our algorithms, we use the 76~GB
BigBrain image split in 125 chunks. The data is available
\todo{[here]}.

\subsubsection{Infrastructure}
 We
perform experiments using the two disks below:
- HD ...
- SSD ...

\todo{Benchmark the disks to measure $\alpha_r$, $\alpha_w$ and $\beta$.}

\subsubsection{Software implementation}

Our algorithms are implemented in Python using the Nibabel
library. \todo{Software release.}

@@

\todo{Valerie's results at PERFORM + model.}

\subsection{Sorted blocks vs blocks}

\subsection{Buffered slices vs slices}

\subsection{Cluster reads vs Multiple reads vs Buffered slices}

\subsection{Model-based comparisons}

Figure~\ref{fig:models} plots the total merge time obtained with the
different algorithms, using the values in~\ref{sec:benchmark} and
different amounts of available memory. Multiple reads behave
consistently better regardless of the amount of available
memory. Cluster reads reach similar performance as available memory
increases.
\begin{figure}
  \includegraphics[width=\columnwidth]{figures/model.pdf}
  % see function.py
  \caption{Model comparison}
  \label{fig:models}
\end{figure}
\todo{redraw the figure. Benchmark $\alpha$ and $\beta$ more
  accurately on the HDD or SSD. In the previous sections, compare the
  model with real experiments, to validate the model.}

\subsection{Lossless Compression}

\subsection{Splitting}
